Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-08-04 volunteer-clouds-ieee-ci-si.tex
Answering @pacastillo questions

There is, then, space for systems that do not have any of those
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. NodIO is first an application that
runs in two tiers, a server and a client that can be possibly run on a  % [Pedro] Both the server and the client in browsers? different browsers as I understand?
browser. As such, it can be run in any computer that loads the web
page where the client is embedded, including possibly many over which
the scientist has no control.

When faced with the problem of diminishing funds for buying new     % [Pedro] This sentence is so so long. I would split it in two sentences.
hardware and continuous hardware failures of existing clusters and
taking into account our experience of working with web applications,
including JavaScript, from the beginning and other volunteer and
unconventional distributed evolutionary computing systems, our group
decided to create a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.

@@

The researcher only has to change a function to solve a different
problem, in this case the classical Trap function \cite{Ackley1987} has been 
used. JavaScript is a functional language and declaring a different
function and handling at the creation of the algorithm object, called %FERGU: poner algo tras el handling: handling it?
{\tt Classic} is the only change needed to work with a different
problem. Let us see how this system addresses the different challenges
outlined in the State of the Art:\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in NodeJS and
  Express.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a single thread
  allows the service of many requests. In fact, a limit with the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \cite{gecco07:workshop:dcor}, which had a low scalability.
\item The system is {\em heterogeneous} since it does not need any
  performance, operating system or even browser requirement: anyone
  visiting the page, even from mobile devices, can load the algorithm. %Fergu: puesto lo de mobile devices para venderlo aún mejor
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved simply through autonomous
  operation of every individual island and no synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server.
\item Since the algorithm is run on the browser, {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources that the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which means that operation is totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL and available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is bigger than what you would obtain in
  your own desktop system. If that is not done, it is a pointless
  academic exercise.
\end{itemize}

>>>>

There is, then, space for systems that do not have any of those
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. NodIO is first an application that
runs in two tiers, a server and a client, the latter  can be possibly
run on a  % [Pedro] Both the server and the client in browsers?
          % different browsers as I understand?  No, server in
          % server, client in browser, although there's nothing that
          % can't be done in a browser. 
browser. As such, it can be run in any computer that loads the web
page where the client is embedded, including possibly many over which
the scientist has no control.

Some time ago, our group faced  the problem of diminishing funds for buying new     % [Pedro] This sentence is so so long. I would split it in two sentences.
hardware and continuous hardware failures of existing
clusters. Considering this, we leveraged our experience of working
with web applications, including JavaScript, since their inception and other volunteer and
unconventional distributed evolutionary computing systems, and we 
decided to designd and relase a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.

@@

The researcher only has to change a function to solve a different
problem, in this case the classical Trap function \cite{Ackley1987} has been 
used. JavaScript is a functional language and declaring a different
function and handling at the creation of the algorithm object, called %FERGU: poner algo tras el handling: handling it?
{\tt Classic} is the only change needed to work with a different
problem. Let us see how this system addresses the different challenges
outlined in the State of the Art (Section \ref{sec:soa}):\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in NodeJS and
  Express.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a single thread
  allows the service of many requests. In fact, a limit with the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \cite{gecco07:workshop:dcor}, which had a low scalability.
\item The system is {\em heterogeneous} since it does not need any
  performance, operating system or even browser requirement: anyone
  visiting the page, even from mobile devices, can load the algorithm. %Fergu: puesto lo de mobile devices para venderlo aún mejor
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved simply through autonomous
  operation of every individual island and no synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server.
\item Since the algorithm is run on the browser, {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources that the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which means that operation is totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL and available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is bigger than what you would obtain in
  your own desktop system. If that is not done, it is a pointless
  academic exercise.
\end{itemize}
